---
title: "Week4"
author: "Jackie Schwartz"
date: "7/24/2019"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
---

The goal for today is to review last week's material on CI and SE bars  
with a new dataset.

# Loading Libraries
```{r}
library(tidyverse)
library(haven)
library(modelr)
library(readxl)
library(NHANES)
library(assertthat)
library(datasets)
```

# Loading dataset
```{r}
tg <- data.frame(datasets::ToothGrowth)
glimpse(tg)
```
What do you notice?  
3 variables:   
`len` (numeric),   
`supp` (factor - OJ or vitamin C),   
`dose` (numeric)
60 observations    

What's our IV?  

What's our DV?  

# Create a summary table  

## Grouping by tx group

```{r}
sum_table <-
  tg %>%
  group_by(supp) %>%
  summarise(
    mean_len = mean(len),
    sd_len = sd(len),
    min_len = min(len),
    max_len = max(len)
  )
```

## Grouping by tx group and dose  
Even though dose is numeric, it really should be treated like an  
ordinal (or factor) so that we can group by it (small, med, large) 

```{r}
tg <-
  tg %>%
  mutate(
    dose_fct =
      as.factor(dose),
    dose_fct =
      recode_factor(dose_fct,
                    "0.5" = "small",
                    "1" = "med",
                    "2" = "lrg")
    )
sum_table <-
  tg %>%
  group_by(supp, dose_fct) %>%
  summarise(
    mean_len = mean(len),
    sd_len = sd(len),
    min_len = min(len),
    max_len = max(len)
  )
```

# What does this look like?

```{r}
nbins <- nclass.FD(tg$len)
rng <- range(tg$len)
bw = (rng[2] - rng[1])/nbins
```

## distribution of tooth length by supp tx
```{r}
g1 <-
  tg %>%
  ggplot(.,
    aes(x = len,
        fill = supp)
  ) +
  geom_histogram(
    bins = nbins,
    position = "dodge",
    color = "black"
  ) +
  scale_fill_manual(values = c("orange", "yellow")) +
  labs(title = "Distribution of tooth growth by supp type") +
  xlab("tooth length") +
  ylab("count") +
  theme_bw()
print(g1)
```

## distribution of tooth length by supp tx and dose
```{r}
g1 <-
  tg %>%
  ggplot(.,
    aes(x = len,
        fill = supp)
  ) +
  geom_histogram(
    bins = nbins,
    position = "dodge",
    color = "black"
  ) +
  facet_wrap(~dose_fct) +
  scale_fill_manual(values = c("orange", "yellow")) +
  labs(title = "Distribution of tooth growth by supp type") +
  xlab("tooth length") +
  ylab("count") +
  theme_bw()
print(g1)
```

__What now?__  
Is there a statitical difference between supplement type in their effect  
on tooth growth? and is this difference _moderated_ by dose?

```{r}
# one approach
mod1 <- t.test(len ~ supp, data = tg)
# second approach
mod2 <- lm(len ~ supp * dose_fct, data = tg)
summary(mod2)
contrasts(tg$supp)
contrasts(tg$dose_fct)
```


```{r}
g2 <-
  tg %>%
  ggplot(.,
    aes(x = supp, y = len, fill = dose_fct)
  ) +
  geom_col(position = "dodge") +
  scale_fill_manual(
    values = c("pink", "royalblue3", "violetred4")
    ) +
  labs(title = "Tooth Growth by Supplement and Dose") +
  xlab("Supplement Type") +
  ylab("Tooth Length") +
  theme_bw()
print(g2)
```

Let's put error bars on this graph!  
SEM = stand dev/sqrt(n)
What is SEM? stand error of the mean
CI(95%) = mean ± 1.96 * SEM  

A 95% confidence interval for a range of scores is
constructed such that the population parameter
will fall within this range in 95% of samples.  

When we want more than just a best guess for a parameter, we can
identify a range of plausible values.

E.g., What’s a reasonable range of patient outcomes a drug could improve?

```{r}
sem_and_ci <-
  tg %>% 
  group_by(supp, dose_fct) %>% 
  summarise(n = n(),
            mean_len = mean(len),
            se_len = sd(len)/sqrt(n),
            ci_lower =
              mean_len - qt(0.975, df = n-1) * se_len,
            ci_upper =
              mean_len + qt(0.975, df = n-1) * se_len
            ) # For each tail covering .025 of the distribution
print(sem_and_ci)
```
## SEM Bars
```{r}
g3 <-
  sem_and_ci %>%
  ggplot(., 
         aes(x = supp, y = mean_len, fill = dose_fct)
         ) +
  geom_col(alpha = 0.5, position = "dodge") +
  geom_errorbar(aes(ymin = mean_len - se_len, 
                    ymax = mean_len + se_len),
                width = .2,
                position = position_dodge(.9)) +
  scale_fill_manual(
    values = c("pink", "royalblue3", "violetred4")
    ) +
  labs(title = "Tooth Growth by Supplement and Dose") +
  xlab("Supplement Type") +
  ylab("Tooth Length") +
  theme_bw()
print(g3)
  
```

## CI bars
```{r}
g4 <-
  sem_and_ci %>%
  ggplot(., 
         aes(x = supp, y = mean_len, fill = dose_fct)
         ) +
  geom_col(alpha = 0.5, position = "dodge") +
  geom_errorbar(aes(ymin = ci_lower, 
                    ymax = ci_upper),
                width = .2,
                position = position_dodge(.9)) +
  scale_fill_manual(
    values = c("pink", "royalblue3", "violetred4")
    ) +
  labs(title = "Tooth Growth by Supplement and Dose") +
  xlab("Supplement Type") +
  ylab("Tooth Length") +
  theme_bw()
print(g4)
```

## Box plot?  
A boxplot splits the data set into quartiles.   
The body of the boxplot consists of a "box" (hence, the name),   
which goes from the first quartile (Q1) to the third quartile (Q3). 

Within the box, a horizontal line is drawn at the Q2,   
the median of the data set, and two vertical lines,   
called whiskers, extend from the top and bottom of the box.   
The bottom whisker goes from Q1 to the smallest non-outlier   
in the data set, and the top whisker goes   
from Q3 to the largest non-outlier.
```{r}
g5 <-
  tg %>%
  ggplot(.,
         aes(x = supp, y = len, fill = dose_fct)) +
  geom_boxplot(alpha = .5, width = 0.3) +
  scale_fill_manual(
    values = c("pink", "royalblue3", "violetred4")
    ) +
  labs(title = "Tooth Growth by Supplement and Dose") +
  xlab("Supplement Type") +
  ylab("Tooth Length") +
  theme_bw()
print(g5)
```

# If we have time...  

Clean this dataset "https://www.kaggle.com/osmi/mental-health-in-tech-survey"  
To click (CMD + click)
```{r}
ment_health_fp <- "/Users/jsschwartz/Dropbox/Stanford_Affective_Science/28b_SummerRtutorial/2019/survey.csv"
ment_health_df <-
  read_csv(ment_health_fp)
```

__What are our variables?__
__Timestamp__  
__Age__  
__Gender__  
__Country__  
__state:__ If you live in the United States,  
which state or territory do you live in? self_employed: Are you self-employed?    
__family_history:__ Do you have a family history of mental illness?  
__treatment:__ Have you sought treatment for a mental health condition?  
__work_interfere:__ If you have a mental health condition,   
do you feel that it interferes with your work?   
__no_employees:__ How many employees does your company or organization have?  
__remote_work:__ Do you work remotely (outside of an office)   
at least 50% of the time?    
__care_options:__ Do you know the options for mental health care   
your employer provides?  
__wellness_program:__ Has your employer ever discussed mental health   
as part of an employee wellness program?  
__seek_help:__ Does your employer provide resources to learn more   
about mental health issues and how to seek help?  
__anonymity:__ Is your anonymity protected if you choose to take   
advantage of mental health or substance abuse treatment resources?  
__leave:__ How easy is it for you to take medical leave  
for a mental health condition?  
__mental_health_consequence:__ Do you think that discussing a mental health  
issue with your employer would have negative consequences?  
__phys_health_consequence:__ Do you think that discussing a physical health  
issue with your employer would have negative consequences?  
__coworkers:__ Would you be willing to discuss a mental health issue  
with your coworkers?  
__supervisor:__ Would you be willing to discuss a mental health issue   
with your direct supervisor(s)?  
__mental_health_interview:__ Would you bring up a mental health issue  
with a potential employer in an interview?  
__phys_health_interview:__ Would you bring up a physical health issue  
with a potential employer in an interview?  
__mental_vs_physical:__ Do you feel that your employer takes mental health  
as seriously as physical health?  
__obs_consequence:__ Have you heard of or observed negative consequences   
for coworkers with mental health conditions in your workplace?  
__comments:__ Any additional notes or comments  

## What about the data class/type of these variables?
```{r}
glimpse(ment_health_df)
```
## converting data types
```{r}
to_factor <-
  function(x) as.factor(x)
ment_health_df <-
  ment_health_df %>%
  mutate_at(vars(-Timestamp,-Age), 
            to_factor) # don't need to do this with converting to factor,  
# numeric, etc... types. But just showing you how you'd do this if you wanted  
# to create a function for other purposes (e.g., scaling variables)
```

Continue cleaning for HW!  
How do plan to get rid of people with unplausable resposnes? (e.g., just  
scanning through the data I found someone with a negative age - yikes!).  
You should also create a unique ID per row. 
Is it useful to combine states to look by region? (NW, SW, MW, NE, SE, MidAt)  
How do you want to deal with the "don't knows?"  
How are you going to check for any weird responses? It's a lot of data  
to manually scan!
```{r}
#maybe convert age to factor just to see what the levels are?
ment_health_df_pos_age <-
  ment_health_df %>%
  mutate(age_f = as.factor(Age)) %>%
  filter(Age > 0 &
           Age < 100) %>%
  mutate(age_f = as.factor(Age))
```

