---
title: "Week1"
author: "Jackie Schwartz"
date: "6/26/2019"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
---

Install these libraries first
# Loading Libraries
```{r}
library(tidyverse)
library(haven)
library(modelr)
library(readxl)
library(NHANES)
library(assertthat)
library(datasets)
```

__To Insert Chunk__  
CMD+OPT+I
```{r}

```

__Basics:__    
_Installing Packages_ 
_Loading Libraries_  
_Importing data_  
__Data Types__  
_Learn R data types of variables_   
__Data Cleaning__  
_Changing variable names_  
_Wide to long format; long to wide format_  
_Changing levels within a factor_ 
_How to replace values (e.g., “999” to NA)_  
_Creating variables (e.g., creating a dummy variable or a “total” variable signifying sum of a scale)_  
_We will be piping throughout_  

# Reading in Data
```{r}
nh_df <- NHANES::NHANES
# in this case package::df

# airquality_df <-
#  datasets::airquality
```

## Ways to view your data
```{r}
# glimpse(nh_df)
# # View(nh_df)
# # nh_df$Poverty
# head(nhanes)
```

Look at num of observations and num of variables. Look at the `class` of  
your variables.
What do you notice about ID?
Do you need to convert any variables?

## Converting data type
```{r}
nhanes_convert <-
  nh_df %>% # piping structure
  mutate( # edits things
    ID_fct = # new variable
      as.factor(ID) # existing variable
         )





nhanes_convert <-
  nh_df %>% 
  mutate(
    ID = as.factor(ID)
    )


```

I'm interested in looking at the association between poor mental health  
and hours of sleep in males and females under 30.  
What variable would represent poor mental health?  Let's use `DaysMentHlthBad`  
What variable would represent hours of sleep? `SleepHrsNight`  
How to "filter" data to those under age 30.

```{r}
nhanes_under_thirty <-
  nhanes_convert %>%
  filter(
    Age < 30
  ) %>%
  select(
    ID,
    Gender,
    Age,
    DaysMentHlthBad,
    SleepHrsNight
    )
```

### how can we confirm that we are only capturing those under 30 yrs old?
```{r}
nhanes_under_thirty <-
  nhanes_under_thirty %>%
  mutate(
    under_thirty =
      Age < 30)
# just to check that our larger dataframe has falses
nh_df <-
  nh_df %>%
  mutate(
    under_thirty =
      Age < 30)
```

Ok, so now we have our under 30 dataframe we are going to work from.  

## add a variable that bins ages  
_changes continuous variable to categorical variable_
0-5  
6-10  
11-15  
15-20  
21-25  
25-30
```{r}
nhanes_clean <-
  nhanes_under_thirty %>%
  mutate(
    age_bin = 
      ifelse(Age < 6,
             "0-5", 
             NA),
    age_bin =
      ifelse(
        Age > 5 &
          Age < 11,
        "6-10",
        age_bin
      ),
    age_bin =
      ifelse(
        Age > 10 & 
          Age < 16,
        "11-15",
        age_bin
      ),
    age_bin =
      ifelse(
        Age > 15 &
          Age < 21,
        "16-20",
        age_bin
      ),
    age_bin =
      ifelse(
        Age > 20 &
          Age < 26,
        "21-25",
        age_bin
      ),
    age_bin =
      ifelse(
        Age > 25 &
          Age < 31,
        "26-30",
        age_bin
      )
  )
### continue with the rest
```
Think about what this if else statement is useful for! 999s 888s
ifelse(variable == "999" | "888",
NA,
variable)
ifelse(variable == "999",
"missing",
variable)
ifelse(variable == "888",
"not_admin",
variable)
## recoding age factor  
_changing levels of a factor_
```{r}
nhanes_clean <-
  nhanes_clean %>%
  mutate(
    age_bin =
      as.factor(age_bin)
  ) %>%
  mutate(age_bin_rec =
           recode_factor(
             age_bin,
             "0-5" = "0", # left side of = is original, right side is new
             "6-10" = "1",
             "11-15" = "2",
             "16-20" = "3",
             "21-25" = "4",
             "26-30" = "5")
         ) %>%
  mutate(gender_rec=
           recode_factor(
             Gender,
             "male" = "0",
             "female" = "1"
           )
         ) %>%
  select(-c(Gender, under_thirty))
```

# 
KNIT!
```{r}
nhanes_reorder <-
  nhanes_clean %>%
  select(
    ID, 
    Age,
    age_bin,
    gender_rec,
    everything()
  )
```


